Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-12-16 analysis/usage.Rmd
minor

```{r variants,echo=F}
nb_calls <- E %>%
  count(wt = duplicates) %>%
  .$n
eval_calls <- E %>%
  filter(ev_variant == "eval") %>%
  count(wt = duplicates) %>%
  .$n
evalq_calls <- E %>%
  filter(ev_variant == "evalq") %>%
  count(wt = duplicates) %>%
  .$n
eval.parent_calls <- E %>%
  filter(ev_variant == "eval.parent") %>%
  count(wt = duplicates) %>%
  .$n
local_calls <- E %>%
  filter(ev_variant == "local") %>%
  count(wt = duplicates) %>%
  .$n

@@

Here we compute the number of static call sistes to eval to compare it with
the number of call sites with triggered dynamically. The former is larger
than the latter because code coverage is not perfect.

@@

TBL <- overview_table(
  # Sites triggered over the whole corpus:
  r("nb eval call sites", sites),
  # How many packages in the corpus have a single triggered call site:
  r("nb one call site", sites_per_package %>% filter(n == 1)),
  r("at least one call site", nrow(sites_per_package)),
  r("packages no calls to eval", corpus_size - nrow(sites_per_package)),
  r("nb median call sites", median(sites_per_package$n))
)
```

How many call sites and how many packages did we exercise out of the total number of call sites of the corpus?

```{r exercised-call-sites, include=F}

# TODO :: remove?
static_sites <- nrow(P)

@@

r("nb static call sites", static_sites)
r("call sites percent", ratio(sites, static_sites))
r("nb static at least one call site", nrow(static_at_least_one_call_site))
r("static at least one call site percent", ratio(nrow(sites_per_package %>% filter(n >= 1)), nrow(static_at_least_one_call_site)))
```

@@

We first compute an approximation of the number of runs per package.
```{r runs-per-package}
runs_package <- E %>%
  group_by(ev_package) %>%
  summarise(nb_runs = n_distinct(file)) %>%
  arrange(desc(nb_runs))
```

@@

weighted.median <- function(x, w, na.rm = FALSE) {
  if (na.rm) {
    na.omit(x)
  }
  w <- w[order(x)]
  x <- x[order(x)]

  prob <- cumsum(w) / sum(w)
  ps <- which(abs(prob - .5) == min(abs(prob - .5)))
  return(x[ps])
}

@@

```{r}
eval_call_triggered <- E %>%
  group_by(src_ref) %>%
  summarize(n_runs = n_distinct(file), n = sum(duplicates), average = n / n_runs) %>%
  arrange(desc(average))
```

```{r}
eval_call_triggered %>% 
  filter(average < 50) %>% 
  ggplot(aes(x = average)) +
  geom_histogram(binwidth =  1)

@@

# The case of core libraries

>>>>

```{r variants, include=FALSE}
nb_calls <- E %>%
  count(wt = duplicates) %>%
  .$n
eval_calls <- E %>%
  filter(ev_variant == "eval") %>%
  count(wt = duplicates) %>%
  .$n
evalq_calls <- E %>%
  filter(ev_variant == "evalq") %>%
  count(wt = duplicates) %>%
  .$n
eval.parent_calls <- E %>%
  filter(ev_variant == "eval.parent") %>%
  count(wt = duplicates) %>%
  .$n
local_calls <- E %>%
  filter(ev_variant == "local") %>%
  count(wt = duplicates) %>%
  .$n

@@

Here we compute the number of static call sites to compare it with
the number of call sites with triggered dynamically. The former is larger
than the latter because code coverage is not perfect.

@@

static_sites <- nrow(P)

@@

overview_table(
  r("Sites", sites),
  r("Staticsites", static_sites),
  r("Sitespercent", ratio(sites, static_sites)),
  r("Staticatleastonecallsite", nrow(static_at_least_one_call_site))
)
```

@@

We first compute an approximation of the number of runs per package.
```{r runs-per-package, include=FALSE}
runs_package <- E %>%
  group_by(ev_package) %>%
  summarise(nb_runs = n_distinct(file)) %>%
  arrange(desc(nb_runs))
```

@@

weighted.median <- function(x, w, na.rm = FALSE) {
  if (na.rm) { na.omit(x) }
  w <- w[order(x)]
  x <- x[order(x)]
  prob <- cumsum(w) / sum(w)
  ps <- which(abs(prob - .5) == min(abs(prob - .5)))
  return(x[ps])
}

@@

```{r}
eval_call_triggered <- E %>%
  group_by(src_ref) %>%
  summarize(n_runs = n_distinct(file), n = sum(duplicates), average = n / n_runs) %>%
  arrange(desc(average))

eval_call_triggered %>% 
  filter(average < 50) %>% 
  ggplot(aes(x = average)) +
  geom_histogram(binwidth =  1)

@@

### 2020-12-16 paper/main.tex
minor

To provide a picture of \eval usage in R, we recorded \Allcalls
invocations of \eval and its variants in the CRAN corpus. These calls
were obtained by \Nbruns runs, with runs that did not trigger \eval
discarded. Out of the \NbStaticAtLeastOneCallSite CRAN packages(among the \Corpus packages of our corpus) with \eval call sites, we
observed calls in \Triggeredpkgs packages, the difference can be chalk
down to incomplete code coverage and occasional failures due to our
pipeline.

@@

\begin{figure}[h]
    \centering
    \begin{tabular}{|r@{\,}r@{\,}lr|r@{\,}r@{\,}lr|} \hline
        \tt 0 &--& \tt 50      & \Runbina  & \tt 501 &--&\tt 1000  & \Runbine\\
        \tt 51 &--& \tt 100    & \Runbinb  & \tt 1001 &--&\tt 1500  & \Runbinf\\
        \tt 101 &--& \tt 250    & \Runbinc  & \tt 1501 &--&\tt 2000   & \Runbing\\
        \tt 251 &--& \tt 500    & \Runbind  & \tt 2001 &--& \tt 3000 & \Runbinh\\\hline
    \end{tabular}
    \caption{Invocations of \eval per call sites normalized by the number of runs.} \label{tab:callsitesnormalized}
\end{figure}

>>>>

To provide a picture of \eval usage in R, we recorded \Allcalls
invocations of \eval and its variants in the CRAN corpus. These calls
were obtained by \Nbruns runs, with runs that did not trigger \eval
discarded. Out of the \Staticatleastonecallsite CRAN packages(among
the \Corpus packages of our corpus) with \eval call sites, we observed
calls in \Triggeredpkgs packages, the difference can be chalk down to
incomplete code coverage and occasional failures due to our pipeline.

@@

\begin{figure}[h] \centering
\begin{tabular}{|r@{\,}r@{\,}lr|r@{\,}r@{\,}lr|} \hline
\tt 0 &--& \tt 50    & \Runbina & \tt 501 &--&\tt 1000   & \Runbine\\
\tt 51 &--& \tt 100  & \Runbinb & \tt 1001 &--&\tt 1500  & \Runbinf\\
\tt 101 &--& \tt 250 & \Runbinc & \tt 1501 &--&\tt 2000  & \Runbing\\
\tt 251 &--& \tt 500 & \Runbind & \tt 2001 &--& \tt 3000 & \Runbinh\\\hline
\end{tabular}
\caption{Normalized invocations} \label{tab:callsitesnormalized}
\end{figure}
